---
title: "recipes"
author: "Mara Alexeev"
date: "10/8/2020"
output: html_document
---
# Part 1: Problem Definition
I love to cook. However, recipe storage is a bit complicated because recipes I use come from so many different sources: cookbooks, emails, online, note cards, printed papers, my memory! For several years, I have wanted to come up with a recipe storage system that can harmonize my recipes and also give me some functionality that is otherwise not present with many formats--scaling, changing units, etc--and that I could eventually turn into a website that I can share with family and friends or just have for myself if traveling. 

Cooking is one thing, but shopping and menu planning are actually the bigger issues for me. Making a single recipe isn't that difficult, but when you start planning a week's worth of cooking or a multi-course meal, the pain points of collecting needed information from so many recipes becomes unwieldy (how many red onions and how many yellow onions do I need to make it through the week?!?)

I actually wrote a [blog post](https://www.maraalexeev.com/2020/04/12/recipe-storage/) about the topic back in April 2020. 

# Part 2: Requirements Gathering

## Users

I imagine my users will be myself, my husband, my brothers, and several close friends. I want it to be available for use to others beyond this circle, but with read-only capacity. 

If the database works very well, I might also offer it to a cousin who runs a resturant to see if he would find a verison for himself to be helpful. 

## Uses

1. Store recipes
2. Read recipes
3. Create menus
4. Aggregate shopping lists from selected recipes
5. Search for recipes by different characteristics
  + Ingredients (eg what's in season, what they have at home already)
  + Holiday favorites
  + Dietary restrictions (eg vegetarian, vegan)
6. Rate recipes
7. Find recipes that pair well with others

## Information Storage

1. Full instructions on how to make the recipe
2. Ingredients needed
3. Equipment needed
4. Ratings from different people over time
5. General information about ingredients
  + When they are in season
  + Where they are typically found in the grocery store
  + What form are they needed for the recipe (eg cubed, diced)

## Updates to database

I anticipate that people will frequently add to recipes to the database but that the tables themselves will not need to be frequently updated. 

## Analytic needs

Several analytic tools I have considered:

1. Looking up recipes that are suitable for guests with different dietary restrictions
2. Looking up some ones rating of a dish before making it again for them
3. Reviewing the notes of people giving ratings to see if there were suggestions for modifications or pairings
4. Making weekly meal planning simpler!
  + What recipes have common features that could allow me to minimize my time in the kitchen? (Eg making more of a particular thing that could be used in multiple meals for the week.)

# Part 3: Database Design

```{r libraries, message=FALSE, warning=FALSE}

library(tidyverse)
library(RSQLite)
library(datamodelr)
library(yaml)
library(DiagrammeR)
```

```{r}
dm <- dm_read_yaml("./recipe_db.yml")
graph <- dm_create_graph(dm, rankdir = "BT")
database_layout <- dm_render_graph(graph)
database_layout

```


# Create tibbles for all tables in database

## Metadata section

### Table: recipe_meta 
```{r}
recipe_meta <- tibble(
  recipe_id = 1:2,
  recipe_name = c("PBJ sandwich", "Srambled eggs"), 
  attribution_id = 1:2, 
  url = c(NA, NA),
  notes = c("Test entry", "Another test"),
  active_time = c(2, 10),
  passive_time = c(0, 2),
  estimated_servings = c(1, NA)
  )

recipe_meta 
```

### Table: menu
```{r}
menu <- tibble(
  recipe_id = 1:2,
  menu_id= 1:2
)

menu
```

### Table: menu_names
```{r}
menu_names <- tibble(
  menu_id = 1:3,
  menu_name = c("Thanksgiving 2020", "New Years Eve 2019", "Birthday Brunch June 2020")
 )

menu_names
```


### Table: courses
```{r}
courses <- tibble(
  recipe_id = 1:2,
  course_id = 1:2
)

courses
```

### Table: course_mapping
```{r}

course_mapping <- tibble(
  course_id = 1:2,
  course_name = c("casual lunch", "breakfast")
  
)

course_mapping 
```

## Feedback section

### Table: users
```{r}
users <- tibble(
  user_id = 1:6,
  name = NA,
  info = NA
)

users 
```

### Table: difficulty
```{r}
difficulty <- tibble(
  difficulty_rating = 1:5,
  description = c("Very Easy", "Easy", "Moderately difficult", "Difficult", "Very Complex")
)

difficulty 
```

### Table: users_notes
```{r}
users_notes <- tibble(
  user_id = 1:2,
  recipe_id = 1:2,
  date = c(20201023, 20200903),
  note = c("My experience was...", "I enjoyed..."),
  proxy_note = as.logical(TRUE, FALSE),
  proxy_id = 3:4
)

users_notes
```

### Table: ratings
```{r}
ratings <- tibble(
  recipe_id = 1:2,
  rating = 1:2,
  rater_id = 1:2,
  difficulty_rating = c(1, NA),
  date = lubridate::ymd(c(20201023, 20200903)),
  notes = c("Terrible", "Would like to try with rye bread toast.")
)

ratings
```


## Ingredients section

### Table: ingredients_used
```{r}
ingredients_used <- tibble(
  recipe_id = 1,
  ingredient_id = 1:4,
  step_number = 1:4,
  quantity = c(2, 3, 4, 5),
  unit = as_factor(c("slices", "cups", "grams", "tablespoons")),
  preparation = as_factor(c("chopped", "diced", "sliced", "minced")),
  notes = NA
)

ingredients_used
```

### Table: ingredients_info
```{r}
ingredients_info <- tibble(
  ingredient_id = 1,
  name = "Carrot",
  season_att = as_factor("Fall"),
  category = as_factor("Vegetable"),
  store_location = as_factor("Produce"),
  ontology_info = "http://purl.obolibrary.org/obo/FOODON_00001687"
)

ingredients_info
```

## Collections section

### Table: tags
```{r}
tags <- tibble(
  recipe_id = 1:2,
  tag_id = 1:2)

tags
```

### Table: tag_names
```{r}
tag_names <- tibble(
  tag_id = 1:4,
  tag_description = as_factor(c("Christmas", "New Years", "Packed Lunch", "Baby Food"))
)

tag_names
```

## Tools and Techniques section

### Table: tools_needed 
```{r}
tools_needed <- tibble(
  recipe_id = 1,
  tool_id = 1
)

tools_needed
```

### Table: tool_info 
```{r}
tool_info <- tibble(
  tool_id = 1:2,
  name = as_factor(c("sous vide", "large pot")),
  cost = c(100, 20),
  alternative = c("water bath", NA),
  notes = NA
)

tool_info 
```

### Table: techniques_used
```{r}
techniques_used <- tibble(
  recipe_id = 1, 
  tech_id = c(1)
)

techniques_used
```


### Table: tech_info
```{r}
tech_info <- tibble(
  tech_id = 1,
  name = "Kneading", 
  notes = "As alternative can use bread setting on mixer."
)

tech_info
```

## Recipe Details section

### Table: steps
```{r}
steps <- tibble(
  recipe_id = 1,
  step_number = 1,
  directions = "Place holder"
)

steps
```

### Table: variations
```{r}
variations <- tibble(
  recipe_id = 1,
  variation_number = 1, 
  variation_text = "Instead of peanut butter try almond butter. It's worse."
)
variations
```
### Table: tips
```{r}
tips <- tibble(
  recipe_id = 1,
  tip_number = 1, 
  tip_text = "Can freeze overnight and will thaw by lunch if put in lunchbag."
)

tips
```

### Table: dietary_modifications
```{r}
dietary_modifications <- tibble(
  recipe_id = 1, 
  modification_id = 1
)

dietary_modifications
```


### Table: dietary_modifications_id
```{r}
dietary_modifications_id <- tibble(
   modification_id = 1,
   mod_id_text = "Vegan Friendly"
)

dietary_modifications_id
```

### Table: abbr_recipes
```{r}
abbr_recipes <- tibble(
  recipe_id = 1,
  step_number = 1,
  abbr_text = "Place holder on abbreviated recipe step."
)

abbr_recipes
```

### Table: raw_recipe_input
```{r}
raw_recipe_input <- tibble(
  recipe_name = "Grandma Donna's Hot Fudge Sauce",
  recipe_id = 3,
  input_text = "Place holder"
)

raw_recipe_input
```


# Create .db files with RSQLite
```{r}
con_recipe_db <- dbConnect(RSQLite::SQLite(), "recipesDB.db")
dbWriteTable(con_recipe_db, "recipe_meta", recipe_meta, overwrite = TRUE)
dbWriteTable(con_recipe_db, "users", users, overwrite = TRUE)
dbWriteTable(con_recipe_db, "users_notes", users_notes, overwrite = TRUE)
dbWriteTable(con_recipe_db, "difficulty", difficulty, overwrite = TRUE)
dbWriteTable(con_recipe_db, "ratings", ratings, overwrite = TRUE)
dbWriteTable(con_recipe_db, "menu", menu, overwrite = TRUE)
dbWriteTable(con_recipe_db, "menu_names", menu_names, overwrite = TRUE)
dbWriteTable(con_recipe_db, "courses", courses, overwrite = TRUE)
dbWriteTable(con_recipe_db, "course_mapping", course_mapping, overwrite = TRUE)
dbWriteTable(con_recipe_db, "steps", steps, overwrite = TRUE)
dbWriteTable(con_recipe_db, "dietary_modifications", dietary_modifications, overwrite = TRUE)
dbWriteTable(con_recipe_db, "dietary_modifications_id", dietary_modifications_id, overwrite = TRUE)
dbWriteTable(con_recipe_db, "abbr_recipes", abbr_recipes, overwrite = TRUE)
dbWriteTable(con_recipe_db, "variations", variations, overwrite = TRUE)
dbWriteTable(con_recipe_db, "tips", tips, overwrite = TRUE)
dbWriteTable(con_recipe_db, "raw_recipe_input", raw_recipe_input, overwrite = TRUE)
dbWriteTable(con_recipe_db, "tags", tags, overwrite = TRUE)
dbWriteTable(con_recipe_db, "tag_names", tag_names, overwrite = TRUE)
dbWriteTable(con_recipe_db, "tools_needed", tools_needed, overwrite = TRUE)
dbWriteTable(con_recipe_db, "tool_info", tool_info, overwrite = TRUE)
dbWriteTable(con_recipe_db, "techniques_used", techniques_used, overwrite = TRUE)
dbWriteTable(con_recipe_db, "tech_info", tech_info, overwrite = TRUE)



dbListTables(con_recipe_db)
```
# Test simple queries with RSQLite
```{r}
dbGetQuery(con_recipe_db, "SELECT * FROM recipe_meta LIMIT 10")
dbGetQuery(con_recipe_db, "SELECT * FROM variations LIMIT 10")
dbGetQuery(con_recipe_db, "SELECT * FROM tag_names LIMIT 10")
```
# Part 4: Ontologies and Terminologies

# Part 5: Analytics

## Question 1

Imagine I am feeling excited about celebrating New Years, but I am stuck in a rut of doing the same recipes for the past few years. 

### Part A

I could search the database to find ingredients that are in season around that time of the year and see if any recipes using them sound good.

### Part B

I am traveling and staying at rental house, but I am going to host a dinner for New Years. I have a restricted number of kitchen equipment, so I want to only consider recipes in my "party" tag that have the equipment i have access to. 
```{r}
dbGetQuery(conn,"SELECT * 
                 FROM recipes
                 LEFT JOIN sources ON source = id_source
                 WHERE source = 1")
```

### Question 2

My parents are coming to visit and but my vegan friend is also coming to dinner. I want to find dishes that my father has rated highly, but are also vegan friendly, so I don't have to make two separate dinners. 

# Bonus: Build it!


## Details
```{r}
sessionInfo()
```

