---
title: "Recipe Database"
author: "Mara Alexeev"
date: "10/8/2020"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

# Part 1: Problem Definition
I love to cook. However, recipe storage is a bit complicated because recipes I use come from so many different sources: cookbooks, emails, online, note cards, printed papers, my memory! For several years, I have wanted to come up with a recipe storage system that can harmonize my recipes and also give me some functionality that is otherwise not present with many formats--scaling, changing units, etc--and that I could eventually turn into a website that I can share with family and friends or just have for myself if traveling. 

Cooking is one thing, but shopping and menu planning are actually the bigger issues for me. Making a single recipe isn't that difficult, but when you start planning a week's worth of cooking or a multi-course meal, the pain points of collecting needed information from so many recipes becomes unwieldy (how many red onions and how many yellow onions do I need to make it through the week?!?)

I actually wrote a [blog post](https://www.maraalexeev.com/2020/04/12/recipe-storage/) about the topic back in April 2020. 

So I have decided to finally take the plunge with this assignment to make the database! I hope that I can use these COVID times of no dinner parties to put my energy into making this a great product that will help me in the years to come. 

# Part 2: Requirements Gathering

## Users

I imagine my users will be myself, my husband, my brothers, and several close friends. I want it to be available for use to others beyond this circle, but with read-only capacity. I plan to post it on github and allow others to fork (yay for food puns!) it for their own uses.

If the database works very well, I might also offer it to a cousin who runs a restaurant to see if he would find a venison for himself to be helpful. 

## Uses

1. Store recipes
2. Read recipes
3. Create menus
4. Aggregate shopping lists from selected recipes
5. Search for recipes by different characteristics
    + Ingredients (eg what's in season, what they have at home already)
    + Holiday favorites
    + Dietary restrictions (eg vegetarian, vegan)
6. Rate recipes
7. Find recipes that pair well with others

## Information Storage

1. Full instructions on how to make the recipe
2. Ingredients needed
3. Equipment needed
4. Ratings from different people over time
5. General information about ingredients
    + When they are in season
    + Where they are typically found in the grocery store
    + What form are they needed for the recipe (eg cubed, diced)

## Updates to database

I anticipate that people will frequently add to recipes to the database but that the tables themselves will not need to be frequently updated. 

## Analytic needs

Several analytic tools I have considered:

1. Looking up recipes that are suitable for guests with different dietary restrictions
2. Looking up some one's rating of a dish before making it again for them
3. Reviewing the notes of people giving ratings to see if there were suggestions for modifications or pairings
4. Making weekly meal planning simpler!
    + What recipes have common features that could allow me to minimize my time in the kitchen? (Eg making more of a particular thing that could be used in multiple meals for the week.)

# Part 3: Database Design

```{r libraries, message=FALSE, warning=FALSE}

library(tidyverse)
library(RSQLite)
library(datamodelr)
library(yaml)
library(DiagrammeR)
library(DiagrammeRsvg)
library(rsvg)
library(readxl)
library(prettydoc)
```

## Database layout
```{r, database layout}
dm_recipe <- dm_read_yaml("./recipe_db.yml")
recipe_graph <- dm_create_graph(dm_recipe, rankdir = "BT")
recipe_database_layout <- dm_render_graph(recipe_graph)
recipe_database_layout

```

# Create tibbles for all tables in database

## Metadata Section

The metadata section of database information about the recipe and how an individual recipe might connect to other recipes to form menus, what courses the recipe is typically served in, and what other recipes are similar to it (eg a seafood paella and a vegetarian paella.)

### Table: recipe_meta 

This table contains the high level information about the recipe. Every recipe should only have a single entry in this form. 

The `recipe_id` column contains an numeric id that uniquely identifies the recipe. The number has no semantic meaning; it is assigned as the recipe is added to the database to the next available integer value. I have considered that a recipe with important variations could be nested under the value. For example, a pasta dish could be 1 and then variations of it could be 1.1, 1.2, etc. However, I decided it would be better for ever recipe to be distinct and instead have variations of recipes linked together by information from the `variations` table. 

The `recipe_name` is the string of text that the recipe is called that would be referred to by people in speech or likely when searching for a recipe.

The `attribution_id` is a numeric value assigned to the creator of the recipe. 

The `recipe_source` is a string that specifies the location if applicable that the recipe can be found at. Examples might be a URL or a book and page number. 

The `recipe_notes` column is where text about the recipe is contained. This might be a history of where the recipe came from or what inspired it, but it does not contain the recipe itself. 

The `active_time` and `passive_time` are time durations that the recipe takes. I had initially thought that this should be stored in minutes, but some recipes have very long passive times, over days, so it might be best to store this in a format that is amenable to being converted to a human friendly display, so that if a recipe has a 3 day passive period it says 3 days not 4320 minutes.

The `estimated_servings` column has the number of servings for the basic recipe. In the shiny app I am building, this will adjust based on the scaling factor input by the user.
```{r recipe_meta table creation}
# recipe_meta <- tibble(
#   recipe_id = 1:2,
#   recipe_name = c("Apple Crisp", "Srambled eggs"), 
#   attribution_id = c(4, 2), 
#   recipe_source = c(NA, NA),
#   recipe_notes = c("A simple and delicious dessert", "Another test"),
#   active_time = c(15, 10),
#   passive_time = c(40, 2),
#   estimated_servings = c(9, NA)
#   )

recipe_meta <-read_excel("files/recipe_workbook.xlsx", 
    sheet = "recipe_meta")

recipe_meta
```

### Table: menu

The `menu` table assign recipes to different menus. A recipe can be assigned many different menus and menus can have many different recipes. Examples of a menu can be a specific event, such as "Mara's 30th Birthday Party" or "New Year's Eve 2018", or to a generic event such as "Brunch" or "Picnic". The overall point of this table versus the tag feature, is that the menu represents curated menus that can be reused in part or in whole.

The `recipe_id` links to the recipes and the `menu_id` will link the name of the menu in the `menu_names` table. The `menu_id` is a numeric value.

```{r}
# menu <- tibble(
#   recipe_id = 1:2,
#   menu_id= c(4, NA)
# )

menu <-read_excel("files/recipe_workbook.xlsx", 
    sheet = "menu")

menu
```

### Table: menu_names

The `menu_names` table provides the mapping between `menu_id` and the human friendly name of the menu in `menu_name`.
```{r}
# menu_names <- tibble(
#   menu_id = 1:4,
#   menu_name = c("Thanksgiving 2020", "New Years Eve 2019", "Birthday Brunch June 2020", "Fall Dinner Party")
#  )

menu_names <-read_excel("files/recipe_workbook.xlsx", 
    sheet = "menu_names")

menu_names
```


### Table: courses

The table `courses` contains information one what courses the recipe is likely to go well in. The `course_id` is a numeric value. A recipe might have multiple courses that it would go well in such as a side dish or as a main entree.
```{r}
# courses <- tibble(
#   recipe_id = 1,
#   course_id = 8
# )

courses <-read_excel("files/recipe_workbook.xlsx", 
    sheet = "courses")

courses

```

### Table: course_mapping

The `course_mapping` table gives the name for the `course_id`. The `course_name` is a string.
```{r}

course_mapping <- tibble(
  course_id = 1:10,
  course_desc = c("Hors-d'oeuvres", "Amuse-gueule", "Soup", "Appetizer", "Salad", "Main Dish", "Palate cleanser", "Dessert", "Drinks", "Snack")

)

course_mapping 
```
### Table: variations
The `variations` table includes pairs of recipes that are closely related to each other have distinct ingredients or steps. For example, a pasta dish might have a meat version and a non meat version but are nearly identical in other respects. 
```{r}
variations <-read_excel("files/recipe_workbook.xlsx", 
    sheet = "variations")

variations
```

## Feedback section

The feedback section of the database includes tables that have to do with people's reactions to recipes or their notes. It also contains the table with information about people in the database. 

### Table: users

The `users`table provides every user for the database with a unique numeric id called user_id. This table is also used to provide names for sources, who might not use the database, but need to be identified else where in the database. An example would be my late grandmother or Bonnie Frumkin Morales, one of my family's favorite chefs whose Russian cookbook, <u>Kachka: A Return To Russian Cooking</u>, is the ground truth for modern Russian cuisine in our home.

The `user_name` column is the person's name stored as a string, and the `user_info` column are additional notes about the user stored as text. 
```{r}
# users <- tibble(
#   user_id = 1:4,
#   user_name = c("Mara Alexeev", "Boris Alexeev", "Denise Determan", "Donna Goodman"),
#   user_info = NA
# )

users <-read_excel("files/recipe_workbook.xlsx", 
    sheet = "users")

users
```

### Table: difficulty

The `difficulty` table is how users can rate recipes. The `recipe_meta` table has a difficulty column, and that is from me as the owner of the database. However, people can keep their own notes and reviews of recipes and use the same scale but choose a different level for the recipe. 
```{r}
difficulty <- tibble(
  difficulty_rating = -2:2,
  difficulty_desc = c("Very Easy", "Easy", "Moderately difficult", "Difficult", "Very Complex")
)

difficulty 
```

### Table: users_notes

The `users_notes` table allows users to store notes about recipes. This is the bug report part of the recipe database. A user can have multiple notes for a recipe. The date allows a person to capture time data about the note, the note field is text. A note can be entered by a proxy then the `proxy_id` is the `user_id` of who did write the note. The use case for this is something like, I write a note for someone, say my young son, who can't write his own notes yet. This would be for ideas about how to change the recipe, or if something needed to be fixed in the recipe. 
```{r}
# users_notes <- tibble(
#   user_id = 1:2,
#   recipe_id = 10:11,
#   note_date = c(20201023, 20200903),
#   user_note = c("Need to update recipe to correct mistake in step 8...", "Too much sauce to noodle ratio..."),
#   proxy_id = 3:4
# )

users_notes <-read_excel("files/recipe_workbook.xlsx", 
    sheet = "users_notes")

users_notes
```

### Table: ratings

The `ratings` table is a way for users to rate recipes on and put in long text reviews. This is different from the `notes` table. The imagined case for this is collect ratings of recipes from both as an indicator of how much someone enjoyed it, to how difficult they found it to make if they cooked. The `try_again` column is a logical value if the person would try the recipe again. The required fields are `recipe_id` and at least one of `rating_id`, `eat_again`, or `difficulty_rating`.
```{r}
# ratings <- tibble(
#   recipe_id = 1:2,
#   rating_id = c(0, 1),
#   try_again = as.logical(c(1, 1)),
#   rater_id = 1:2,
#   difficulty_rating = c(0, NA),
#   rating_date = lubridate::ymd(c(20201023, 20200903)),
#   rating_note = c("Wonderful!", "Would like to try with rye bread toast."),
#   proxy_id = c(1,4)
# )

ratings <-read_excel("files/recipe_workbook.xlsx", 
    sheet = "ratings")

ratings
```

### Table: rating_scale

The `rating_scale` table maps the numeric value of the `rating_id` to a text factor in `rating_desc` that describes the numeric values's meanings. 

The `rating` scale is a minus 2 to plus 2 scale which is highly favored in the Alexeev household. It has the benefit that the values 1 and 5 can be confused (which is the good value??), but that people are unlikely to think -2 is a good value to assign to a recipe. 

```{r}
rating_scale <- tibble(
  rating_id = -2:2,
  rating_desc = c("Strongly disliked", "Disliked", "Indifferent", "Enjoyed", "Strongly Enjoyed")
)

rating_scale
```


## Ingredients section
The `ingredients` section of the database has tables with information about ingredients needed for recipes and then also gives information about individual ingredients unrelated to any specific recipe such as where the ingredient is likely to be found in the grocery store. 
### Table: ingredients_used
```{r}
# ingredients_used <- tibble(
#   recipe_id = 1,
#   ingredient_id = c(2, 3, 4, 5),
#   step_number = 1:4,
#   quantity = c(2, 3, 4, 5),
#   unit = as_factor(c("slices", "cups", "grams", "tablespoons")),
#   scaling_factor = 1,
#   preparation = as_factor(c("chopped", "diced", "sliced", "minced")),
#   ingredient_note = NA
# )



ingredients_used <- read_excel("files/recipe_workbook.xlsx", 
    sheet = "ingredients_used")

ingredients_used

```

### Table: ingredients_info
The `ingredients_info` table links ingredients to ontology information for the ingredients. The onology is called "FoodOn" and is discussed in [Part 4]. 
```{r}
# ingredients_info <- tibble(
#   ingredient_id = c(1, 2, 3, 4, 5, 6),
#   ingredient_name = c("Carrot", "Apple, Gold Delicious", "Sugar", "Cinnamon", "Butter", "Flour"), 
#   season_attr = as_factor(c("Fall", "Fall", "None", "None", "None", "None")),
#   category = as_factor(c("Vegetable", "Fruit", "Dry good", "Spice", "Dairy", "Dry Good")),
#   store_location = c("Produce", "Produce", "Baking", "Spice", "Dairy", "Baking"),
#   ontology_info = c("http://purl.obolibrary.org/obo/FOODON_00001687", "http://www.ontobee.org/ontology/FOODON?iri=http://purl.obolibrary.org/obo/FOODON_03411245", "http://www.ontobee.org/ontology/FOODON?iri=http://purl.obolibrary.org/obo/CHEBI_17992", "http://www.ontobee.org/ontology/FOODON?iri=http://purl.obolibrary.org/obo/FOODON_03301556", "http://www.ontobee.org/ontology/FOODON?iri=http://purl.obolibrary.org/obo/FOODON_03301777", "http://www.ontobee.org/ontology/FOODON?iri=http://purl.obolibrary.org/obo/FOODON_03304534")
# )

ingredients_info <- read_excel("files/recipe_workbook.xlsx", 
    sheet = "ingredients_info")

ingredients_info
```

## Collections section
The collections section allows users to tag recipes with a concept to allow personalized grouping. A recipe can have multiple tags. Tags can represent a concept like time of year the dish is often associated such as Thanksgiving, or it can tag the recipe with an idea like that it is a good dish for packed lunches. This also can tag the recipe with a concept like the type of cuisine it is, eg Russian.
### Table: tags
```{r}
# tags <- tibble(
#   recipe_id = 1,
#   tag_id = 6)

tags <-  read_excel("files/recipe_workbook.xlsx", 
    sheet = "tags")

tags
```

### Table: tag_names
```{r}
# tag_names <- tibble(
#   tag_id = 1:6,
#   tag_desc = as_factor(c("Christmas", "New Years", "Packed Lunch", "Baby Food", "Russian", "Goodman Family Tradition"))
# )

tag_names <-  read_excel("files/recipe_workbook.xlsx", 
    sheet = "tag_names")

tag_names
```

## Tools and Techniques section

These tables include information about what is required to make the recipes excluding the ingredients. The `tools` table and `tool_info` tables include things like knives, equipment, types of pans, etc needed to make the recipe and some information on substitutions. 
### Table: tools_needed 
```{r}
# tools_needed <- tibble(
#   recipe_id = 1,
#   tool_id = 1
# )

tools_needed <-  read_excel("files/recipe_workbook.xlsx", 
    sheet = "tools_needed")

tools_needed
```

### Table: tool_info 
```{r}
# tool_info <- tibble(
#   tool_id = 1:7,
#   tool_name = as_factor(c("chef's knife", "peeler", "measuring spoons", "measuring cups", "oven", "sous vide", "large pot")),
#   tool_cost = NA,
#   tool_alternative = NA,
#   tool_note = NA
# )

tool_info <-  read_excel("files/recipe_workbook.xlsx", 
    sheet = "tool_info")

tool_info
```

### Table: techniques_used

The `techniques_used` and `tech_info` tables include information about what special techniques, excluding things like washing, cutting, peeling, that are needed for the recipe. These become important for high difficulty recipes that the novice cook would not be familiar with. 
```{r}
# techniques_used <- tibble(
#   recipe_id = 2, 
#   tech_id = c(1)
# )

techniques_used <-  read_excel("files/recipe_workbook.xlsx", 
    sheet = "techniques_used")

techniques_used
```


### Table: tech_info
```{r}
# tech_info <- tibble(
#   tech_id = 1,
#   tech_name = "Kneading", 
#   tech_note = "As alternative can use bread setting on mixer."
# )

tech_info <-  read_excel("files/recipe_workbook.xlsx", 
    sheet = "tech_info")

tech_info
```

## Recipe Details section
The recipes details section contains 3 tables that include the information on the precise steps to make a recipe. 

### Table: steps

The `steps` table includes the directions to create the recipe. The `version_id` replaces a different table that was originally included as a separate table that included abbreviated versions of the recipe. As I was building the database I realized it was easier to have a recipe with different versions in this table and not need to worry about updating other tables in the ingredients section and metadata section. The Apple Crisp Recipe is an example of this. Note that ingredients are bolded for help when following directions. Also note that the recipe gives ratios of ingredients. I pondered this decision for multiple days before deciding on ratios instead of absolute values to allow for easier development later as I build up the scaling option in the Shiny app. 
```{r}
# steps <- tibble(
#   recipe_id = 1,
#   step_number = c(1, 2, 3, 4),
#   directions = c("Peel and slice **all apples** and place in baking dish.", "Sprinkle **half of the sugar** and **2/3 of the cinnamon** over the apples. Mix to coat apples evenly.", "In a separate bowl, mix the other **half of the sugar**, **1/3 of the cinnamon**, **all the flour**, and **all the butter**. Sprinkle mixture over apples.", "Bake at 350F degrees for 40 minutes. The topping should be lightly browned.")
# )

steps <-  read_excel("files/recipe_workbook.xlsx", 
    sheet = "steps")

head(steps)

```

### Table: tips
The `tips` table includes tips about recipes. The inclusion of this table in the next version is under debate as it seems like tips and user_notes could be combined perhaps with a note type indicating what the note was about (eg tip, bug, to do.)
```{r}
# tips <- tibble(
#   recipe_id = 99,
#   tip_number = 1, 
#   tip_text = "Can freeze overnight and will thaw by lunch if put in lunchbag."
# )

tips <- read_excel("files/recipe_workbook.xlsx", 
    sheet = "tips")

tips
```



### Table: raw_recipe_input
The `raw_recipe_input` is the quick and dirty way to get a recipe into the system for future processing for full integration into the system. This is something my mother could do via the shiny app that doesn't require special knowledge of the database and is easy to submit to me for inclusion in the database. 
```{r}
# raw_recipe_input <- tibble(
#   recipe_name = "Grandma Donna’s Chocolate Sauce",
#   recipe_id = 2,
#   raw_text = "1/4 pound butter (one stick), 3 squares (3 oz) chocolate (this would be 12 now, as Bakers’ Chocolate comes in ¼ oz “squares”...used to be 1 oz squares. Just be mindful of whatever the brand is packaged as...that is, by weight.), 3 cups sugar, Dash salt, 1 tall can evaporated milk (Carnation Milk)
# Melt butter and chocolate over hot water. (I use a double boiler.) Stir in sugar one tablespoon at a time. (This is tedious, but important to get smooth results.) Add milk and heat until sauce is velvety smooth and sugar has all been dissolved. When adding sugar, DON’T Panic! The mixture becomes VERY thick and sugar isn’t dissolving. Try to get it all added, if possible. Then add the milk. I use a whisk to stir. By the time it is done, no sugar crystals should remain."
# )

raw_recipe_input <- read_excel("files/recipe_workbook.xlsx", 
    sheet = "raw_recipe_input")

raw_recipe_input


```


# Create .db files with RSQLite
This is how I built the database with the RSQlite package. 
```{r}
con_recipe_db <- dbConnect(RSQLite::SQLite(), "recipesDB.db")
dbWriteTable(con_recipe_db, "recipe_meta", recipe_meta, overwrite = TRUE)
dbWriteTable(con_recipe_db, "users", users, overwrite = TRUE)
dbWriteTable(con_recipe_db, "users_notes", users_notes, overwrite = TRUE)
dbWriteTable(con_recipe_db, "difficulty", difficulty, overwrite = TRUE)
dbWriteTable(con_recipe_db, "ratings", ratings, overwrite = TRUE)
dbWriteTable(con_recipe_db, "menu", menu, overwrite = TRUE)
dbWriteTable(con_recipe_db, "menu_names", menu_names, overwrite = TRUE)
dbWriteTable(con_recipe_db, "courses", courses, overwrite = TRUE)
dbWriteTable(con_recipe_db, "course_mapping", course_mapping, overwrite = TRUE)
dbWriteTable(con_recipe_db, "steps", steps, overwrite = TRUE)
dbWriteTable(con_recipe_db, "ingredients_used", ingredients_used, overwrite = TRUE)
dbWriteTable(con_recipe_db, "ingredients_info", ingredients_info, overwrite = TRUE)
dbWriteTable(con_recipe_db, "variations", variations, overwrite = TRUE)
dbWriteTable(con_recipe_db, "tips", tips, overwrite = TRUE)
dbWriteTable(con_recipe_db, "raw_recipe_input", raw_recipe_input, overwrite = TRUE)
dbWriteTable(con_recipe_db, "tags", tags, overwrite = TRUE)
dbWriteTable(con_recipe_db, "tag_names", tag_names, overwrite = TRUE)
dbWriteTable(con_recipe_db, "tools_needed", tools_needed, overwrite = TRUE)
dbWriteTable(con_recipe_db, "tool_info", tool_info, overwrite = TRUE)
dbWriteTable(con_recipe_db, "techniques_used", techniques_used, overwrite = TRUE)
dbWriteTable(con_recipe_db, "tech_info", tech_info, overwrite = TRUE)



dbListTables(con_recipe_db)

```
# Test simple queries with RSQLite
```{r, results= 'asis'}
test_1 <- dbGetQuery(con_recipe_db, "SELECT * FROM recipe_meta LIMIT 10")


knitr::kable(test_1)
```
```{r, results= 'asis'}
test_2 <-  dbGetQuery(con_recipe_db, "SELECT * FROM ingredients_used LIMIT 10")

knitr::kable(test_2)
```

```{r, results= 'asis'}
test_3 <-  dbGetQuery(con_recipe_db, "SELECT * FROM ingredients_info LIMIT 10")

knitr::kable(test_3)
```

```{r, results= 'asis'}
test_4 <- dbGetQuery(con_recipe_db, "SELECT * FROM recipe_meta WHERE recipe_name LIKE '%salad%'")

knitr::kable(test_4)
```

Note the difference between of the recipes versions.
```{r full vs abbr recipes, results= 'asis'}
recipe_comparison <- dbGetQuery(con_recipe_db, 
"SELECT recipe_name, steps.recipe_id, version_id, step_number, directions
FROM steps
LEFT JOIN  recipe_meta ON steps.recipe_id=recipe_meta.recipe_id
Where recipe_meta.recipe_id = 1
ORDER by steps.recipe_id, version_id")

knitr::kable(recipe_comparison)
```

# Part 4: Ontologies and Terminologies

• When, where and how you will use the terminology or ontology in your database?

My ingredients table contains a field with the url linking to the FoodOn database. The future use for the link would be to link to outside the database potentially to a food purchasing entity that would allow for automating purchasing. Additionally, common names of some foods in English are often regional and do not map to the same item to people (eg yams and sweet potatoes, eggplant and aubergine. The FoodOn information could provide an unambiguous link to the item and also might be able to pull images of the food in a future iteration of the project. 

• Who develops the terminology or ontology?

FoodOn continues to be developed by an open consortium of university researchers. It was started in 2015 by the [Hsiao Lab from University of British Columbia](https://foodon.org/about/). 

• How is it maintained?

The project is open source and has a [github repo](https://github.com/FoodOntology/foodon/issues) that people can contribute to and is currently quite active. Several research groups continue to develop it and it is funded by several universities and several governmental agencies from Canada. 

• How well does it meet Cimino’s desiderata?
Overall the ontology seems to be very informed on the desiderata and build off the structures of other ontologies. 

Content: Yes, the ontology is exanpansive and can be reused in a variety of areas. It continues to expand to meet new needs. The expansion is guided by developed rules and is regulated by a group.

Nonvaguness/nonambiguity: Yes 

Code permanence: Yes. Terms can be deprecated but they are not removed. The deprecated term has links to point a user to the new terms. 

Nonsematic identifers: True.

Polyhierarchy: Yes.

Formal Definitions: Yes.

Granularities: Yes. There is apple as term but then different cultivares have their own subterm from apple. 

Multiple Consistent Views: Yes

Representing Context: Yes the whole point of the ontology is to put food in context from every part of its lifecycle to how it is grown and harvested to how it is transformed in preservation and cooking.

Evolve Gracefully: Yes! This is a great strength of the project that is very transparent about its changes and has active discussions on its github account on how to adapt the ontology as it grows. 

Recognize Redundancy: This is the least clear. I could not find an example of this, but I think it is likely after I read the curation section of the ontology that discussed how terms might be coded differently but mean the same thing and should map to the same idea (boneless vs deboned.) 

• Why did you pick it?

It was the most comprehensive and actively developed ontology in the area that I could find.

• Were there alternatives?

I could find no viable alternatives.

# Part 5: Analytics

## Question 1

I am making two recipes and I want to make a combined shopping list with ingredients, but if ingredients are shared, I want to have them combined into one place on the list, eg 2 onions, not 1 onion listed two times on the list. 
```{r, results = 'asis'}
q1 <- dbGetQuery(con_recipe_db, 
"SELECT ingredient_name, sum(quantity), unit, store_location
FROM ingredients_used
LEFT JOIN  ingredients_info ON ingredients_used.ingredient_id=ingredients_info.ingredient_id
Where recipe_id = 1 OR 2
group by ingredient_name, unit
ORDER by store_location

")

knitr::kable(q1)

```


## Question 2

I hosting a family reuion for my mother's side of the family, so I want to make some recipes from her mother, Donna Goodman. I also want to display which courses those recipes fall under, so I can see which other places I will need to round out my menu with recipes, perhaps from other Goodman family members.
```{r}
q2 <- dbGetQuery(con_recipe_db,"SELECT recipe_name, user_name AS creator, recipe_notes, course_desc
                 FROM recipe_meta
                 Left join users on recipe_meta.attribution_id = users.user_id
                 LEFT JOIN courses on recipe_meta.recipe_id = courses.recipe_id
                 LEFT JOIN course_mapping ON courses.course_id = course_mapping.course_id
                 WHERE attribution_id = 4")

knitr::kable(q2)
```


### Question 3

My parents are coming to visit and but my vegan friend is also coming to dinner. I want to find dishes that my father has rated highly, but are also vegan friendly, so I don't have to make two separate dinners. Currently the database doesn't have any recipes that fit this because I don't have ratings from my father yet, even when I do, this might be a difficult dinner party to prepare for!! To be continued!

# Bonus: Build it!

App.R and Global.R make up a working skeleton of an app to query and view the tables in relation to each other. I have plans to expand it to allow guided addition of recipes to write to the database and also an automated database query to make shopping lists. 

## Github Repo
[https://github.com/MaraAlexeev/701_database_project_MaraAlexeev](701_database_project_MaraAlexeev)

## Details
```{r}
sessionInfo()
```

